A Stochastic Quasi-Newton Optimizer for TensorFlow

### 2018-05-06 paper/main.tex
Section 2 draft

Wang et al. \cite{sdlbfgs} assume that Problem (\ref{prb:mini}) satisfies the
following assumptions:
\begin{enumerate}
  \item There is a real lower bound for $f$. That is, $\inf_{x \in \RR^n} f(x) >
    -\infty$.
  \item The gradient $\nabla f$ is Lipschitz continuous. That is, there exists
    $L > 0$ such that 
    \begin{equation*}
      \norm{\nabla f(x) - \nabla f(y)} \leq L \norm{x - y}
    \end{equation*}
    for all $x,y \in \RR^n$. Moreover, $\nabla_{xx} F(x, \xi)$ exists and is
    continuous and there exists $\kappa > 0$ with $\norm{\nabla^2_{xx} F(x,
    \varepsilon)} \leq \kappa$ for any $x, \xi$.
  \item The sampling regime is unbiased. That is, for the $k$th iterate the
    following hold:
    \begin{align*}
      \ex{g(x_k, \xi_k)}_{\xi_k} &= \nabla f(x_k), \\
      \ex{\norm{g(x_k, \xi_k) - \nabla f(x_k)}^2}_{\xi_k} &\leq \sigma^2,
    \end{align*}
    where $g(x, \xi_k) = \nabla F(x, \xi_k)$ and $\sigma^2$ is the noise level
    of estimating $g(x, \xi_k)$. 
\end{enumerate}

@@

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SQN}{$x_1, \ip{m_k}_{k=1}^{\infty},
    \ip{\alpha_k}_{k=1}^{\infty}, p$}
    \State{$\textrm{data} \leftarrow \emptyset$}
    \For {$k = 1, 2, \dots$}
    \State{$g_k \leftarrow \frac{1}{m_k} \sum_{i=1}^{m_k} g(x_k,
    \varepsilon_{k,i})$} 
    \State{$\Delta x_k, s_{k-1}, \bar{y}_{k-1}, \rho_{k-1} \leftarrow
    \textsc{SdLBFGS}(g_k, {}^*(\textrm{data}\trans))$} \Comment{Tuple unpacking}
    \State{$x_{k+1} \leftarrow x_k - \alpha_k \Delta x_k$}
    \State{$\textit{append}(\textrm{data}, (s_{k-1}, \bar{y}_{k-1}, \rho_{k-1}))$}
    \If{$k > p$}
    \State{$\textit{pop}(\textrm{data})$}
    \EndIf
    \EndFor
    \EndFunction
  \end{algorithmic}
  \label{alg:sqn}
  \caption{%
    The high level stochastic quasi-Newton minimization algorithm. Given an
    initial point $x_1 \in \RR^n$, batch sizes $\ip{m_k}$, step sizes
    $\ip{\alpha_k}$, and a maximum memory capacity $p$, perform BFGS with the
    stochastic dampened direction update.
  }
\end{algorithm}

>>>>

Wang et al. \cite{sdlbfgs} assume that Problem (\ref{prb:mini}) satisfies the
following assumptions:
\begin{enumerate}
  \item There is a real lower bound for $f$. That is, $\inf_{x \in \RR^n} f(x) >
    -\infty$.
  \item The gradient $\nabla f$ is Lipschitz continuous. That is, there exists
    $L > 0$ such that 
    \begin{equation*}
      \norm{\nabla f(x) - \nabla f(y)} \leq L \norm{x - y}
    \end{equation*}
    for all $x,y \in \RR^n$. Moreover, $\nabla_{xx} F(x, \xi)$ exists and is
    continuous, and there exists $\kappa > 0$ with $\norm{\nabla^2_{xx} F(x,
    \varepsilon)} \leq \kappa$ for any $x, \xi$.
  \item The sampling regime is unbiased. That is, for the $k$th iterate the
    following hold:
    \begin{align*}
      \ex{g(x_k, \xi_k)}_{\xi_k} &= \nabla f(x_k), \\
      \ex{\norm{g(x_k, \xi_k) - \nabla f(x_k)}^2}_{\xi_k} &\leq \sigma^2,
    \end{align*}
    where $g(x, \xi_k) = \nabla F(x, \xi_k)$ and $\sigma^2$ is the noise level
    of estimating $g(x, \xi_k)$. 
\end{enumerate}

@@

We note that the formulation of Algorithm \ref{alg:sdlbfgs} differs slightly
from Wang et al. In the original paper, the use of $\textrm{mem}$ in the section
corresponding to lines 15 and 16 of Algorithm \ref{alg:sdlbfgs} is replaced by
$p$. This leads to the difficulty that if $p$ indicates memory \emph{capacity}
rather than simply the current amount of data in memory, the indices can become
negative early in the iteration. Specifically, when $k < p$ in the original
paper the algorithm has undefined behavior. We clarify this ambiguity in the
following way: in our implementation, $p$ indicates the total memory capacity of
the procedure, regardless of the total amount of memory presently in use. The
quantity $\textrm{mem}$ replaces $p$ in the SdLBFGS update step so as to prevent
array index out of bounds errors when $k$ is small (i.e., $k < p$), but both our
implementation and the original paper's exhibit the same behavior whenever $k >
p$.

\begin{algorithm}
  \begin{algorithmic}[1]
    \Function{SQN}{$x_1, \ip{m_k}_{k=1}^{\infty},
    \ip{\alpha_k}_{k=1}^{\infty}, p$}
    \State{$\textrm{data} \leftarrow \textit{nil}()$}
    \For {$k = 1, 2, \dots$}
    \State{$g_k \leftarrow \frac{1}{m_k} \sum_{i=1}^{m_k} g(x_k,
    \varepsilon_{k,i})$} 
    \State{$\Delta x_k, s_{k-1}, \bar{y}_{k-1}, \rho_{k-1} \leftarrow
    \textsc{SdLBFGS}(g_k, {}^*(\textrm{data}\trans))$} \Comment{Tuple unpacking}
    \State{$x_{k+1} \leftarrow x_k - \alpha_k \Delta x_k$}
    \State{$\textit{append}(\textrm{data}, (s_{k-1}, \bar{y}_{k-1}, \rho_{k-1}))$}
    \If{$k > p$}
    \State{$\textit{pop}(\textrm{data})$}
    \EndIf
    \EndFor
    \EndFunction
  \end{algorithmic}
  \label{alg:sqn}
  \caption{%
    The high level stochastic quasi-Newton minimization algorithm. Given an
    initial point $x_1 \in \RR^n$, batch sizes $\ip{m_k}$, step sizes
    $\ip{\alpha_k}$, and a maximum memory capacity $p$, perform BFGS with the
    stochastic dampened direction update.
  }
\end{algorithm}
